---
title: "Chronic Absenteeism - Trend analysis"
author: "Felicia Zhang"
date: '2018-06-03'
output: pdf_document
toc: yes
toc_depth: 2
fontsize: 12pt
fig_height: 6
fig_width: 7
---

```{r setup, include=FALSE, warning=FALSE}
library(ggplot2) 
library(zoo)
library(reshape)
library(plyr)
library(dplyr)
library(scales) 
library(data.table)
library(signal)
library(matrixStats)
library(lme4)
library(arm)
library(broom)
library(tidyr)
library(knitr)
library(wesanderson)
library(readxl)    
library(extrafont)

#LOAD MULTIPLE EXCEL SHEETS INTO SEPARATE DFS

#get names of sheets
sheets <- readxl::excel_sheets("~/Documents/Felicia Zhang/Felicia/Princeton/ConsultingClub/ChronicAbsenteeism/chronicabsenteeismtrend.xlsx")

#load in excel data as 1 big file
lst <- lapply(sheets, function(sheet) 
  readxl::read_excel("~/Documents/Felicia Zhang/Felicia/Princeton/ConsultingClub/ChronicAbsenteeism/chronicabsenteeismtrend.xlsx", sheet = sheet)
)

#prepare names
names(lst) <- sheets

#turn it into DF
list2env(lst, envir = .GlobalEnv)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
#COMBINE SHEETS INTO ONE DF
poofinal <- rbind(`01_02CA`,`02_03CA`,`03_04CA`,`04_05CA`,`05_06CA`,`06_07CA`,`07_08CA`,`08_09CA`,`09_10CA`,`10_11CA`,`12_13CA`,`13_14CA`,`14_15CA`,`15_16CA`,`16_17CA`)
poofinal2 <- gsub('[%]','',poofinal$PERCENT_CHRONIC_ABSENT)
poofinal$PERCENT_CHRONIC_ABSENT <- poofinal2

#label missing values as NA
poofinal[poofinal == "N"] <- NA
poofinal[poofinal == "*"] <- NA

#convert all columns to numeric
poofinal[] <- lapply(poofinal, function(x) as.numeric(as.character(x)))

#only mercer
mercer <- subset(poofinal, COUNTY_CODE==21)
mercer$YEAR <- factor(mercer$YEAR)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
# create individual DFs
#trend plot
mercer$PERCENT_CHRONIC_ABSENT <- mercer$PERCENT_CHRONIC_ABSENT / 100

#89 different schools, lets break it up into the separate districts
#District 1245 = EASTW = East Windsor
eastwind <- subset(mercer, DISTRICT_CODE==1245)

#District 1430 = EWI = Ewing
ewing <- subset(mercer, DISTRICT_CODE==1430)

#District 1950 = HAM = Hamilton
hamilton <- subset(mercer, DISTRICT_CODE==1950)

#District 2280 = HOP = Hopewell
hopewell <- subset(mercer, DISTRICT_CODE==2280)

#District 2580 = LAW = Lawrence
lawrence <- subset(mercer, DISTRICT_CODE==2580)

#District 3105 = MER = Mercer County Vocational
mercervo <- subset(mercer, DISTRICT_CODE==3105)

#District 3103 = MERSS = Mercer County Special Service
mercerss <- subset(mercer, DISTRICT_CODE==3103)

#District 4255 = PRI = Princeton
princeton <- subset(mercer, DISTRICT_CODE==4255)

#District 5210 = TRE = Trenton
trenton <- subset(mercer, DISTRICT_CODE==5210)

#District 5510 = ROB = Robbinsville
robbins <- subset(mercer, DISTRICT_CODE==5510)

#District 5715 = WPL = Windsor-Plainsboro
winplain <- subset(mercer, DISTRICT_CODE==5715)
```

Important notes:  
1. We do not have data for 11/12 school year  
2. For the school years 10/11 and earlier, we do not chronic absenteeism rate. Instead, we have attendance, so chronic absenteeism rate was calculated using 1 - attendance %.  

\newpage

## Trend analysis for Mercer County
```{r, echo=FALSE, warning=FALSE}
#group schools in each district together
mercertotal <- group_by(mercer, DISTRICT_CODE, YEAR) %>%
  summarise(
    meanPercentabsent = mean(PERCENT_CHRONIC_ABSENT, na.rm = TRUE),
    sePercentabsent=sd(PERCENT_CHRONIC_ABSENT, na.rm = TRUE)/sqrt(length(PERCENT_CHRONIC_ABSENT))
  )

ggplot(mercertotal,aes(x=YEAR,y=meanPercentabsent,group=DISTRICT_CODE, color=factor(DISTRICT_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nMercer County")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.3),breaks=seq(0,.3,.05))+
  scale_color_discrete(name="District",breaks=c("1245","1430","1950","2280","2580","3103","3105","4255","5210","5510","5715"),labels=c("East Windsor", "Ewing","Hamilton","Hopewell", "Lawrence","Special Service","Vocational","Princeton","Trenton", "Robbinsville","Windsor-Plainsboro"))+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")
```
\newpage

## Trend analysis for each district in Mercer County

### East Windsor
```{r, echo=FALSE, warning=FALSE}
# East windsor
ggplot(eastwind,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nEast Windsor")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.16),breaks=seq(0,.16,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Ewing

```{r, echo=FALSE, warning=FALSE}
# Ewing
ggplot(ewing,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nEwing")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.16),breaks=seq(0,.16,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Hamilton
```{r, echo=FALSE, warning=FALSE,fig.height=7}
# hamilton
ggplot(hamilton,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nHamilton")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.26),breaks=seq(0,.26,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Hopewell
```{r, echo=FALSE, warning=FALSE}
# hopewell
ggplot(hopewell,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nHopewell")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.2),breaks=seq(0,.2,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")
```
\newpage

### Lawrence
```{r, echo=FALSE, warning=FALSE}
# lawrence
ggplot(lawrence,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nLawrence")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.2),breaks=seq(0,.2,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Mercer County Vocational
```{r, echo=FALSE, warning=FALSE}
# mercer vocational
ggplot(mercervo,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nMercer County Vocational")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.3),breaks=seq(0,.3,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Mercer County Special Service
```{r, echo=FALSE, warning=FALSE}
# mercer special service
ggplot(mercerss,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nMercer County Special Service")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.2),breaks=seq(0,.2,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Princeton

```{r, echo=FALSE, warning=FALSE,fig.height=6}
# princeton
ggplot(princeton,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nPrinceton")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.32),breaks=seq(0,.32,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Trenton
```{r, echo=FALSE, warning=FALSE,fig.height=6}
# trenton
ggplot(trenton,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nTrenton")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.8),breaks=seq(0,.8,.1))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Robbinsville
```{r, echo=FALSE, warning=FALSE}
# robbins
ggplot(robbins,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nRobbinsville")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.1),breaks=seq(0,.1,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

### Windsor-Plainsboro
```{r, echo=FALSE, warning=FALSE}
# Windsor-Plainsboro
ggplot(winplain,aes(x=YEAR,y=PERCENT_CHRONIC_ABSENT,group=SCHOOL_CODE, color=factor(SCHOOL_CODE)))+
  geom_line(size=2)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nWindsor-Plainsboro")+
  labs(x = "School year", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_discrete(breaks=c("20012002","20022003","20032004","20042005","20052006","20062007","20072008","20082009","20092010","20102011","20122013","20132014","20142015","20152016","20162017"),labels=c("01/02", "02/03","03/04","04/05","05/06","06/07","07/08","08/09","09/10", "10/11","12/13","13/14","14/15","15/16","16/17"))+
  scale_y_continuous(labels=percent,limits=c(0,.2),breaks=seq(0,.2,.02))+
  scale_color_discrete(name="School")+
  theme(text=element_text(family="Times"))+
  theme(axis.text.x = element_text(angle = 90, hjust = 1))+
  theme(legend.position = "bottom")

```
\newpage

## Summary

East Windsor - School 60 increased in 2012/2013

