---
title: "Chronic Absenteeism"
author: "Felicia Zhang"
date: '2018-05-15'
output:
html_document: default
---

```{r setup, include=FALSE, warning=FALSE}
library(ggplot2) 
library(zoo)
library(reshape)
library(plyr)
library(dplyr)
library(scales) 
library(data.table)
library(signal)
library(matrixStats)
library(lme4)
library(arm)
library(broom)
library(tidyr)
library(wesanderson)
library(readxl)    
library(extrafont)

#LOAD MULTIPLE EXCEL SHEETS INTO SEPARATE DFS

#get names of sheets
sheets <- readxl::excel_sheets("~/Documents/Felicia Zhang/Felicia/Princeton/ConsultingClub/PerformanceReportsSchoolData_SY1617_FZ.xlsx")

#load in excel data as 1 big file
lst <- lapply(sheets, function(sheet) 
  readxl::read_excel("~/Documents/Felicia Zhang/Felicia/Princeton/ConsultingClub/PerformanceReportsSchoolData_SY1617_FZ.xlsx", sheet = sheet)
)

#prepare names
names(lst) <- sheets

#turn it into DF
list2env(lst, envir = .GlobalEnv)
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
###Create Chronic Absenteeism percent for each grade, each school, and each district

#*SchoolPercent is n of students absent/n of students in the grade NOT n of students absent/n of students in the school -_-

poo <- EnrollmentTrendsByGrade

#remove ungraded and Total Enrollment
poo <- poo[!(poo$Grade=="Ungraded"),]
poo <- poo[!(poo$Grade=="Total Enrollment"),]

#remove the word Grade from beginning
poo$Grade <- substring(poo$Grade, 7)

#reformat poo Grade (it's written as 07 instead of 7)
poo$Grade[poo$Grade=="01"] <- "1"
poo$Grade[poo$Grade=="02"] <- "2"
poo$Grade[poo$Grade=="03"] <- "3"
poo$Grade[poo$Grade=="04"] <- "4"
poo$Grade[poo$Grade=="05"] <- "5"
poo$Grade[poo$Grade=="06"] <- "6"
poo$Grade[poo$Grade=="07"] <- "7"
poo$Grade[poo$Grade=="08"] <- "8"
poo$Grade[poo$Grade=="09"] <- "9"

poo2 <- ChronicAbsByGrade

#combine the 2
AbsentGrade <- merge(poo, poo2, by=c("CountyCode","DistrictCode","SchoolCode","Grade")) # NA's match

#relabel SchoolPercent to GradePercent
names(AbsentGrade)[5]<-"NumofStudents"
names(AbsentGrade)[6]<-"GradePercent"

#replace grades with numeric
AbsentGrade$Grade[AbsentGrade$Grade=="PK"] <- "-1"
AbsentGrade$Grade[AbsentGrade$Grade=="KG"] <- "0"

#label missing values as NA
AbsentGrade[AbsentGrade == "N"] <- NA
AbsentGrade[AbsentGrade == "*"] <- NA

#convert all columns to numeric
AbsentGrade[] <- lapply(AbsentGrade, function(x) as.numeric(as.character(x)))

#convert GradePercent to decimal
AbsentGrade$GradePercent <- AbsentGrade$GradePercent/100

#calculate Number of students absent using total number of students * percent absent
AbsentGrade$NumStudentsAbsent <- AbsentGrade$NumofStudents * AbsentGrade$GradePercent
AbsentGrade$NumStudentsAbsent <- round(AbsentGrade$NumStudentsAbsent)

#get percent absent for each school
AbsentSchool <- group_by(AbsentGrade,CountyCode,DistrictCode,SchoolCode) %>%
  summarise(
    percentabsent = sum(NumStudentsAbsent) / sum(NumofStudents)
  )

#get percent absent for each district
AbsentDistrict <- group_by(AbsentGrade,CountyCode,DistrictCode) %>%
  summarise(
    percentabsent = sum(NumStudentsAbsent) / sum(NumofStudents)
  )
```

### To get a sense of the data, lets look at the distribution of chronic absenteeism rate
```{r, echo=FALSE, warning=FALSE,message=FALSE,fig.width=7, fig.height=8}
#histogram of school absenteeism rate 
#to see min, max, what’s the most common rate
#color by county
poo <- na.omit(AbsentSchool)

ggplot(poo, aes(percentabsent, fill=factor(CountyCode)))+
  geom_histogram(binwidth = 0.01)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nHistogram of school chronic absenteeism rate")+
  labs(x = "Chronic absenteeism rate", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(labels=percent,limits=c(-0.1,1),breaks=seq(-0.1,1,.1))+
  theme(legend.position = "bottom")+
  theme(text=element_text(family="Century"))+
  guides(fill=guide_legend(title="County"))
```

```{r, echo=FALSE, warning=FALSE}
#Chronic absenteeism rate by grade
poo <- AbsentGrade

poo2 <- group_by(poo, CountyCode, DistrictCode, Grade) %>%
  summarise(
    percentabsent = mean(GradePercent)
  )

poo.final <- group_by(poo2, Grade) %>%
  summarise(
    meanPercentabsent = mean(percentabsent,  na.rm = TRUE),
    sePercentabsent=sd(percentabsent, na.rm = TRUE)/sqrt(length(percentabsent))
  )

limits <- aes(ymax = meanPercentabsent + sePercentabsent, ymin=meanPercentabsent - sePercentabsent)

ggplot(poo.final,aes(x=factor(Grade),y=meanPercentabsent,color=factor(Grade),fill=factor(Grade)))+
  geom_bar(stat="identity")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nChronic absenteeism rate by grade")+
  labs(x = "Grade", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(labels=percent,limits=c(0,.3),breaks=seq(0,.3,.05))+
  scale_x_discrete(breaks=c("-1","0","1","2","3","4","5","6","7","8","9","10","11","12"),labels=c("PK", "KG","1","2","3","4","5","6","7","8","9","10","11","12"))+
  theme(legend.position = "none")+
  geom_errorbar(limits, width=0.25,color="black")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(text=element_text(family="Century"))
```

###Focus on the good schools: which county are they in?
```{r, echo=FALSE, warning=FALSE}
#calculate mean and SD
poo <- na.omit(AbsentSchool)
z <- mean(poo$percentabsent) #0.0976
zz <- sd(poo$percentabsent) #0.0792

#look at good schools: anything less than mean
poo2 <- subset(poo, percentabsent < z)

#plot
ggplot(poo2, aes(percentabsent, fill=factor(CountyCode)))+
  geom_histogram(binwidth = 0.01)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nHistogram of school chronic absenteeism rate (low CA)")+
  labs(x = "Chronic absenteeism rate", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_x_continuous(labels=percent,limits=c(-0.02,.12),breaks=seq(-0.02,.12,.02))+
  theme(legend.position = "bottom")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(text=element_text(family="Century"))+
  guides(fill=guide_legend(title="County"))

#plot histogram of school
poo3 <- group_by(poo2, CountyCode) %>%
  summarise(
    count = n()
  )

#sort by descending order
poo3$CountyCode <- factor(poo3$CountyCode, levels = poo3$CountyCode[order(-poo3$count)])

ggplot(poo3, aes(x=factor(CountyCode),y=count,color=factor(CountyCode),fill=factor(CountyCode)))+
  geom_bar(stat="identity")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nCounty 3 has most schools w low chronic absenteeism")+
  labs(x = "County", y = "Count of schools")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "none")+
  theme(text=element_text(family="Century"))
```

###Focus on the bad schools: which county are they in?
```{r, echo=FALSE, warning=FALSE}
#calculate mean and SD
z <- mean(poo$percentabsent) #0.0976
zz <- sd(poo$percentabsent) #0.0792

#define bad schools: > 2SD from mean 
outliersbad <- z + (2*zz) #anything greater than 25% 

#look at bad schools
poo2 <- subset(poo, percentabsent > outliersbad)

#plot
ggplot(poo2, aes(percentabsent, fill=factor(CountyCode)))+
  geom_histogram(binwidth = 0.01)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nHistogram of high CA (2SD above mean)")+
  labs(x = "Chronic absenteeism rate", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  scale_x_continuous(labels=percent,limits=c(.2,1),breaks=seq(.2,1,.1))+
  theme(legend.position = "bottom")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(text=element_text(family="Century"))+
  guides(fill=guide_legend(title="County"))

#plot histogram of school
poo3 <- group_by(poo2, CountyCode) %>%
  summarise(
    count = n()
  )

#sort by descending order
poo3$CountyCode <- factor(poo3$CountyCode, levels = poo3$CountyCode[order(-poo3$count)])

ggplot(poo3, aes(x=factor(CountyCode),y=count,color=factor(CountyCode),fill=factor(CountyCode)))+
  geom_bar(stat="identity")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nCounty 13 has most schools w high chronic absenteeism")+
  labs(x = "County", y = "Count of schools")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "none")+
  theme(text=element_text(family="Century"))
```

###What are the characteristics of county 13 (high CA) and county 3 (low CA)?
```{r, echo=FALSE, warning=FALSE}
foo <- subset(EnrollmentTrendsByGrade, CountyCode == "13" | CountyCode == "03")

#remove total enrollment
foo <- foo[!(foo$Grade=="Total Enrollment"),]

#remove the word Grade from beginning
foo$Grade <- substring(foo$Grade, 7)

#reformat foo Grade (it's written as 07 instead of 7)
foo$Grade[foo$Grade=="01"] <- "1"
foo$Grade[foo$Grade=="02"] <- "2"
foo$Grade[foo$Grade=="03"] <- "3"
foo$Grade[foo$Grade=="04"] <- "4"
foo$Grade[foo$Grade=="05"] <- "5"
foo$Grade[foo$Grade=="06"] <- "6"
foo$Grade[foo$Grade=="07"] <- "7"
foo$Grade[foo$Grade=="08"] <- "8"
foo$Grade[foo$Grade=="09"] <- "9"
foo$Grade[foo$Grade=="PK"] <- "-1"
foo$Grade[foo$Grade=="KG"] <- "0"
foo$Grade[foo$Grade=="ed"] <- "99" #ungraded?

#label missing values as NA
foo[foo == "N"] <- NA
foo[foo == "*"] <- NA

#convert all columns to numeric
foo[] <- lapply(foo, function(x) as.numeric(as.character(x)))

#calculate total number of students in each grade 
foo2 <- group_by(foo, CountyCode) %>%
  summarise(
    numofstudents = sum(Count)
  )

ggplot(foo2,aes(x=factor(CountyCode),y=numofstudents,color=factor(CountyCode),fill=factor(CountyCode)))+
  geom_bar(stat="identity", position ="dodge")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nNumber of students")+
  labs(x = "County", y = "Number of students")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(legend.position = "none")+
  theme(text=element_text(family="Century"))+
  scale_fill_manual(values = wes_palette("GrandBudapest2"))+
  scale_color_manual(values = wes_palette("GrandBudapest2"))+
  scale_y_continuous(labels = comma, limits=c(0,130000),breaks=seq(0,130000,25000))

```

```{r, echo=FALSE, warning=FALSE,fig.width=8}
foo <- subset(EnrollmentTrendsByGrade, CountyCode == "13" | CountyCode == "03")

#remove total enrollment
foo <- foo[!(foo$Grade=="Total Enrollment"),]

#remove the word Grade from beginning
foo$Grade <- substring(foo$Grade, 7)

#reformat foo Grade (it's written as 07 instead of 7)
foo$Grade[foo$Grade=="01"] <- "1"
foo$Grade[foo$Grade=="02"] <- "2"
foo$Grade[foo$Grade=="03"] <- "3"
foo$Grade[foo$Grade=="04"] <- "4"
foo$Grade[foo$Grade=="05"] <- "5"
foo$Grade[foo$Grade=="06"] <- "6"
foo$Grade[foo$Grade=="07"] <- "7"
foo$Grade[foo$Grade=="08"] <- "8"
foo$Grade[foo$Grade=="09"] <- "9"
foo$Grade[foo$Grade=="PK"] <- "-1"
foo$Grade[foo$Grade=="KG"] <- "0"
foo$Grade[foo$Grade=="ed"] <- "99" #ungraded?

#label missing values as NA
foo[foo == "N"] <- NA
foo[foo == "*"] <- NA

#convert all columns to numeric
foo[] <- lapply(foo, function(x) as.numeric(as.character(x)))

#calculate total number of students in each grade 
foo2 <- group_by(foo, CountyCode, Grade) %>%
  summarise(
    numofstudents = sum(Count)
  )

#calculate total num of students for county 13 and 3
x13 <- sum(foo2$numofstudents[foo2$CountyCode==13])
x3 <- sum(foo2$numofstudents[foo2$CountyCode==3])

#divide to get percent
foo2$percent <- 0
foo2$percent[1:15] <- foo2$numofstudents[1:15] / x3
foo2$percent[16:30] <- foo2$numofstudents[16:30] / x13

ggplot(foo2,aes(x=factor(Grade),y=percent,color=factor(CountyCode),fill=factor(CountyCode)))+
  geom_bar(stat="identity", position ="dodge")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nGrade breakdown")+
  labs(x = "Grade", y = "Percent of students")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(labels=percent,limits=c(0,.1),breaks=seq(0,.1,.01))+
  scale_x_discrete(breaks=c("-1","0","1","2","3","4","5","6","7","8","9","10","11","12","99"),labels=c("PK", "KG","1","2","3","4","5","6","7","8","9","10","11","12","Ungraded"))+
  theme(legend.position = "bottom")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(text=element_text(family="Century"))+
  scale_fill_manual(values = wes_palette("GrandBudapest2"),name="County Code",breaks=c("3","13"),labels=c("3 (lower CA)", "13 (higher CA)"))+
  scale_color_manual(values = wes_palette("GrandBudapest2"))+
  guides(color=FALSE)

```

```{r, echo=FALSE, warning=FALSE, fig.width=10}
foo <- subset(EnrollmentByRacialEthnicGroup, CountyCode == "13" | CountyCode == "03")

#label missing values as NA
foo[foo == "N"] <- NA
foo[foo == "*"] <- NA

#convert all columns to numeric
foo$CountyCode <- as.numeric(as.character(foo$CountyCode))
foo$DistrictCode <- as.numeric(as.character(foo$DistrictCode))
foo$SchoolCode <- as.numeric(as.character(foo$SchoolCode))
foo$Percent <- as.numeric(as.character(foo$Percent))

#convert race group to numbers
foo$ethnic <- 1
foo$ethnic[foo$RacialAndEthnicGroup=="Asian"] <- 2
foo$ethnic[foo$RacialAndEthnicGroup=="Black or African American"] <- 3 
foo$ethnic[foo$RacialAndEthnicGroup=="Hispanic"] <- 4 
foo$ethnic[foo$RacialAndEthnicGroup=="Native Hawaiian or Pacific Islander"] <- 5 
foo$ethnic[foo$RacialAndEthnicGroup=="Two or More Races"] <- 6 
foo$ethnic[foo$RacialAndEthnicGroup=="White"] <- 7 

#convert number to decimal
foo$Percent <- foo$Percent / 100

#calculate total number of students in each grade 
foo2 <- group_by(foo, CountyCode, ethnic) %>%
  summarise(
    meanpercent = mean(Percent),
    sepercent=sd(Percent, na.rm = TRUE)/sqrt(length(Percent))
  )

limits <- aes(ymax = meanpercent + sepercent, ymin=meanpercent - sepercent)

ggplot(foo2,aes(x=factor(ethnic),y=meanpercent,color=factor(CountyCode),fill=factor(CountyCode)))+
  geom_bar(stat="identity", position ="dodge")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nEthnicity breakdown")+
  labs(x = "Grade", y = "Percent of students")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(labels=percent,limits=c(0,.6),breaks=seq(0,.6,.1))+
  scale_x_discrete(breaks=c("1","2","3","4","5","6","7"),labels=c("Native American", "Asian","Black","Hispanic","Hawaiian/PacIslander","2+ Races","White"))+
  theme(legend.position = "bottom")+
  theme(text=element_text(family="Century"))+
  scale_fill_manual(values = wes_palette("GrandBudapest2"),name="County Code",breaks=c("3","13"),labels=c("3 (lower CA)", "13 (higher CA)"))+
  scale_color_manual(values = wes_palette("GrandBudapest2"))+
  geom_errorbar(limits, width=0.25, color="black",position=position_dodge(width=0.9))+
  guides(color=FALSE)
  
```

```{r, echo=FALSE, warning=FALSE,fig.width=10}
foo <- subset(EnrollmentTrendsByStudentGroup, CountyCode == "13" | CountyCode == "03")
foo <- subset(foo, StudentGroup != "Male" & StudentGroup != "Female")

#label missing values as NA
foo[foo == "N"] <- NA
foo[foo == "*"] <- NA

#convert all columns to numeric
foo$CountyCode <- as.numeric(as.character(foo$CountyCode))
foo$DistrictCode <- as.numeric(as.character(foo$DistrictCode))
foo$SchoolCode <- as.numeric(as.character(foo$SchoolCode))
foo$Percent <- as.numeric(as.character(foo$Percent))

#convert race group to numbers
foo$group <- 1
foo$group[foo$StudentGroup=="English Learners"] <- 2
foo$group[foo$StudentGroup=="Homeless Students"] <- 3 
foo$group[foo$StudentGroup=="Migrant Students"] <- 4 
foo$group[foo$StudentGroup=="Military-Connected Students"] <- 5 
foo$group[foo$StudentGroup=="Students in Foster Care"] <- 6 
foo$group[foo$StudentGroup=="Students with Disabilities"] <- 7 

#convert number to decimal
foo$Percent <- foo$Percent / 100

#calculate total number of students in each grade 
foo2 <- group_by(foo, CountyCode, group) %>%
  summarise(
    meanpercent = mean(Percent),
    sepercent=sd(Percent, na.rm = TRUE)/sqrt(length(Percent))
  )

limits <- aes(ymax = meanpercent + sepercent, ymin=meanpercent - sepercent)

ggplot(foo2,aes(x=factor(group),y=meanpercent,color=factor(CountyCode),fill=factor(CountyCode)))+
  geom_bar(stat="identity", position ="dodge")+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  ggtitle("\n\nStudent group breakdown")+
  labs(x = "Grade", y = "Percent of students")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(plot.title = element_text(hjust = 0.5))+
  scale_y_continuous(labels=percent,limits=c(0,.6),breaks=seq(0,.6,.1))+
  scale_x_discrete(breaks=c("1","2","3","4","5","6","7"),labels=c("EcoDis", "EngLearners","Homeless","Migrant","Military","Foster Care","Disabilities"))+
  theme(legend.position = "bottom")+
  theme(text=element_text(family="Century"))+
  scale_fill_manual(values = wes_palette("GrandBudapest2"),name="County Code",breaks=c("3","13"),labels=c("3 (lower CA)", "13 (higher CA)"))+
  scale_color_manual(values = wes_palette("GrandBudapest2"))+
  geom_errorbar(limits, width=0.25, color="black",position=position_dodge(width=0.9))+
  guides(color=FALSE)
```

####Comparing County 3 and County 13, it looks like County 13 (higher CA) have more economically disadvantaged students, less Asian students, less White students, and more Black students. County 13 also tend to have more students in elementary school compared to middle/high school.

##### Next: what are the characteristics of good vs bad schools?
