---
title: "Chronic Absenteeism - Subgroup analysis"
author: "Felicia Zhang"
date: '2018-07-03'
output: pdf_document
toc: yes
toc_depth: 2
fontsize: 12pt
fig_height: 6
fig_width: 7
---

```{r setup, include=FALSE, warning=FALSE}
library(ggplot2) 
library(zoo)
library(reshape)
library(plyr)
library(dplyr)
library(scales) 
library(data.table)
library(signal)
library(matrixStats)
library(lme4)
library(arm)
library(broom)
library(tidyr)
library(knitr)
library(wesanderson)
library(readxl)    
library(extrafont)

#LOAD MULTIPLE EXCEL SHEETS INTO SEPARATE DFS

#get names of sheets
sheets <- readxl::excel_sheets("~/Documents/Felicia Zhang/Felicia/Princeton/ConsultingClub/ChronicAbsenteeism/20162017_SchoolData.xlsx")

#load in excel data as 1 big file
lst <- lapply(sheets, function(sheet) 
  readxl::read_excel("~/Documents/Felicia Zhang/Felicia/Princeton/ConsultingClub/ChronicAbsenteeism/20162017_SchoolData.xlsx", sheet = sheet)
)

#prepare names
names(lst) <- sheets

#turn it into DF
list2env(lst, envir = .GlobalEnv)

#subgroup analysis: using the demographics, show distributions, ratio bw whites and blacks, anything to look at inequality
```

```{r, echo=FALSE, warning=FALSE, include=FALSE}
###Create Chronic Absenteeism percent for each grade, each school, and each district

#*SchoolPercent is n of students absent/n of students in the grade NOT n of students absent/n of students in the school -_-

poo <- ChronicAbsenteeism

#keep only schoolwide data
poo <- subset(poo, StudentGroup=="Schoolwide")

#create school data
AbsentSchool <- poo[1:4]

#label missing values as NA
AbsentSchool[AbsentSchool == "N"] <- NA
AbsentSchool[AbsentSchool == "*"] <- NA

#convert all columns to numeric
AbsentSchool[] <- lapply(AbsentSchool, function(x) as.numeric(as.character(x)))

#convert GradePercent to decimal
AbsentSchool$percentabsent <- AbsentSchool$SchoolPercent/100

#get percent absent for each district
AbsentDistrict <- group_by(AbsentSchool,CountyCode,DistrictCode) %>%
  summarise(
    percentabsent = mean(percentabsent, na.rm = TRUE)
  )

#Mercer county = 21
Mercer <- subset(AbsentSchool, CountyCode==21)
```

Important note:  
All graphs are based on 2016-2017 school year

\newpage
##Correlations 

###Chronic absenteeism and racial groups
```{r, echo=FALSE, warning=FALSE,fig.width=10}
#sort schools by race (Black, White, Hispanic, Asian)
poo <- subset(EnrollmentByRacialEthnicGroup, CountyCode=="21") 
poo$CountyCode <- as.numeric(as.character(poo$CountyCode))
poo$DistrictCode <- as.numeric(as.character(poo$DistrictCode))
poo$SchoolCode <- as.numeric(as.character(poo$SchoolCode))
poo$RacePercent <- as.numeric(as.character(poo$Percent))

poofinal <- merge(poo, Mercer, by=c("CountyCode","DistrictCode","SchoolCode"))

poofinal$SchoolPercent <- poofinal$SchoolPercent /100
poofinal$RacePercent <- poofinal$RacePercent /100

poofinal <- na.omit(poofinal)

ggplot(poofinal,aes(x=RacePercent,y=SchoolPercent))+
  geom_point(size=4)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "none")+
  facet_wrap(~RacialAndEthnicGroup)+
  theme(strip.text = element_text(size=14))+
  stat_smooth(method=lm, se = FALSE)+
  scale_x_continuous(labels=percent)+
  scale_y_continuous(labels=percent)

#only focus on Asian, African American, White, Hispanic
poofinal2 <- subset(poofinal, RacialAndEthnicGroup== "Asian" | RacialAndEthnicGroup=="Black or African American" | RacialAndEthnicGroup== "Hispanic" | RacialAndEthnicGroup== "White")

ggplot(poofinal2,aes(x=RacePercent,y=SchoolPercent))+
  geom_point(size=4)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "none")+
  facet_wrap(~RacialAndEthnicGroup)+
  theme(strip.text = element_text(size=14))+
  stat_smooth(method=lm, se = FALSE)+
  scale_x_continuous(labels=percent)+
  scale_y_continuous(labels=percent)
```
Chronic absenteeism is correlated with percentage of Black/African American students in a school compared to any other race.

\newpage

###Chronic absenteeism and SES groups
```{r, echo=FALSE, warning=FALSE,fig.width=10}
#sort schools by SES
poo <- subset(EnrollmentTrendsByStudentGroup, CountyCode=="21") 
poo$CountyCode <- as.numeric(as.character(poo$CountyCode))
poo$DistrictCode <- as.numeric(as.character(poo$DistrictCode))
poo$SchoolCode <- as.numeric(as.character(poo$SchoolCode))
poo$SESPercent <- as.numeric(as.character(poo$Percent))

poofinal <- merge(poo, Mercer, by=c("CountyCode","DistrictCode","SchoolCode"))

poofinal$SchoolPercent <- poofinal$SchoolPercent /100
poofinal$SESPercent <- poofinal$SESPercent /100

poofinal <- na.omit(poofinal)

ggplot(poofinal,aes(x=SESPercent,y=SchoolPercent))+
  geom_point(size=4)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "none")+
  facet_wrap(~StudentGroup)+
  theme(strip.text = element_text(size=14))+
  stat_smooth(method=lm, se = FALSE)+
  scale_x_continuous(labels=percent)+
  scale_y_continuous(labels=percent)

#only focus on 
poofinal2 <- subset(poofinal, StudentGroup== "Economically Disadvantaged Students" | StudentGroup=="English Learners")

ggplot(poofinal2,aes(x=SESPercent,y=SchoolPercent))+
  geom_point(size=4)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "none")+
  facet_wrap(~StudentGroup)+
  theme(strip.text = element_text(size=14))+
  stat_smooth(method=lm, se = FALSE)+
  scale_x_continuous(labels=percent)+
  scale_y_continuous(labels=percent)

```
Chronic absenteeism is correlated with percentage of low SES and English Learners in a school.

\newpage

###Chronic absenteeism and home language
```{r, echo=FALSE, warning=FALSE,fig.width=10}
#sort schools by language
poo <- subset(EnrollmentByHomeLanguage, CountyCode=="21") 
poo$CountyCode <- as.numeric(as.character(poo$CountyCode))
poo$DistrictCode <- as.numeric(as.character(poo$DistrictCode))
poo$SchoolCode <- as.numeric(as.character(poo$SchoolCode))
poo$LangPercent <- as.numeric(as.character(poo$`%OfStudents`))

poofinal <- merge(poo, Mercer, by=c("CountyCode","DistrictCode","SchoolCode"))

poofinal$SchoolPercent <- poofinal$SchoolPercent /100
poofinal$LangPercent <- poofinal$LangPercent /100

poofinal <- na.omit(poofinal)

#only focus on English, Spanish
poofinal2 <- subset(poofinal, HomeLanguage== "English" | HomeLanguage=="Spanish")

ggplot(poofinal2,aes(x=LangPercent,y=SchoolPercent))+
  geom_point(size=4)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Chronic absenteeism rate")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "none")+
  facet_wrap(~HomeLanguage)+
  theme(strip.text = element_text(size=14))+
  stat_smooth(method=lm, se = FALSE)+
  scale_x_continuous(labels=percent)+
  scale_y_continuous(labels=percent)

```
Chronic absenteeism is correlated with speaking Spanish at home.

\newpage

##Distributions 

###Chronic absenteeism and racial groups
```{r, echo=FALSE, warning=FALSE,fig.width=10}
#sort schools by race (Black, White, Hispanic, Asian)
poo <- subset(EnrollmentByRacialEthnicGroup, CountyCode=="21") 
poo$CountyCode <- as.numeric(as.character(poo$CountyCode))
poo$DistrictCode <- as.numeric(as.character(poo$DistrictCode))
poo$SchoolCode <- as.numeric(as.character(poo$SchoolCode))
poo$RacePercent <- as.numeric(as.character(poo$Percent))

poofinal <- merge(poo, Mercer, by=c("CountyCode","DistrictCode","SchoolCode"))

poofinal$SchoolPercent <- poofinal$SchoolPercent /100
poofinal$RacePercent <- poofinal$RacePercent /100

poofinal <- na.omit(poofinal)

poofinal$CA <- 0
poofinal$CA[poofinal$SchoolPercent > 0.1] <- 1

ggplot(poofinal, aes(RacePercent, fill = factor(CA)))+
  geom_histogram(binwidth=.05)+
  facet_wrap(~RacialAndEthnicGroup)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "bottom")+
  theme(strip.text = element_text(size=14))+
  scale_x_continuous(labels=percent)+
  scale_fill_manual(values = c("#bebad9", "#3b6dae"), name="Chronic absenteeism", breaks=c("0","1"), labels=c("No", "Yes"))

#only focus on Asian, African American, White, Hispanic
poofinal2 <- subset(poofinal, RacialAndEthnicGroup== "Asian" | RacialAndEthnicGroup=="Black or African American" | RacialAndEthnicGroup== "Hispanic" | RacialAndEthnicGroup== "White")

ggplot(poofinal2, aes(RacePercent, fill = factor(CA)))+
  geom_histogram(binwidth=.05)+
  facet_wrap(~RacialAndEthnicGroup)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "bottom")+
  theme(strip.text = element_text(size=14))+
  scale_x_continuous(labels=percent)+
  scale_fill_manual(values = c("#bebad9", "#3b6dae"), name="Chronic absenteeism", breaks=c("0","1"), labels=c("No", "Yes"))
```

\newpage

###Chronic absenteeism and SES groups
```{r, echo=FALSE, warning=FALSE,fig.width=10}
#sort schools by SES
poo <- subset(EnrollmentTrendsByStudentGroup, CountyCode=="21") 
poo$CountyCode <- as.numeric(as.character(poo$CountyCode))
poo$DistrictCode <- as.numeric(as.character(poo$DistrictCode))
poo$SchoolCode <- as.numeric(as.character(poo$SchoolCode))
poo$SESPercent <- as.numeric(as.character(poo$Percent))

poofinal <- merge(poo, Mercer, by=c("CountyCode","DistrictCode","SchoolCode"))

poofinal$SchoolPercent <- poofinal$SchoolPercent /100
poofinal$SESPercent <- poofinal$SESPercent /100

poofinal <- na.omit(poofinal)

poofinal$CA <- 0
poofinal$CA[poofinal$SchoolPercent > 0.1] <- 1

ggplot(poofinal, aes(SESPercent, fill = factor(CA)))+
  geom_histogram(binwidth=.05)+
  facet_wrap(~StudentGroup)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "bottom")+
  theme(strip.text = element_text(size=14))+
  scale_x_continuous(labels=percent)+
  scale_fill_manual(values = c("#bebad9", "#3b6dae"), name="Chronic absenteeism", breaks=c("0","1"), labels=c("No", "Yes"))

#only focus on
poofinal2 <- subset(poofinal, StudentGroup== "Economically Disadvantaged Students" | StudentGroup=="English Learners")

ggplot(poofinal2, aes(SESPercent, fill = factor(CA)))+
  geom_histogram(binwidth=.05)+
  facet_wrap(~StudentGroup)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "bottom")+
  theme(strip.text = element_text(size=14))+
  scale_x_continuous(labels=percent)+
  scale_fill_manual(values = c("#bebad9", "#3b6dae"), name="Chronic absenteeism", breaks=c("0","1"), labels=c("No", "Yes"))
```

\newpage

###Chronic absenteeism and home language
```{r, echo=FALSE, warning=FALSE,fig.width=10}
#sort schools by language
poo <- subset(EnrollmentByHomeLanguage, CountyCode=="21") 
poo$CountyCode <- as.numeric(as.character(poo$CountyCode))
poo$DistrictCode <- as.numeric(as.character(poo$DistrictCode))
poo$SchoolCode <- as.numeric(as.character(poo$SchoolCode))
poo$LangPercent <- as.numeric(as.character(poo$`%OfStudents`))

poofinal <- merge(poo, Mercer, by=c("CountyCode","DistrictCode","SchoolCode"))

poofinal$SchoolPercent <- poofinal$SchoolPercent /100
poofinal$LangPercent <- poofinal$LangPercent /100

poofinal$CA <- 0
poofinal$CA[poofinal$SchoolPercent > 0.1] <- 1

#only focus on English, Spanish
poofinal2 <- subset(poofinal, HomeLanguage== "English" | HomeLanguage=="Spanish")

ggplot(poofinal, aes(LangPercent, fill = factor(CA)))+
  geom_histogram(binwidth=.05)+
  facet_wrap(~HomeLanguage)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=12),axis.text.y  = element_text(size=12))+
  theme(legend.text=element_text(size=12),legend.title=element_text(size=12))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "bottom")+
  theme(strip.text = element_text(size=12))+
  scale_x_continuous(labels=percent)+
  scale_fill_manual(values = c("#bebad9", "#3b6dae"), name="Chronic absenteeism", breaks=c("0","1"), labels=c("No", "Yes"))

ggplot(poofinal2, aes(LangPercent, fill = factor(CA)))+
  geom_histogram(binwidth=.05)+
  facet_wrap(~HomeLanguage)+
  theme_bw()+
  theme(panel.grid.major = element_blank(),panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))+
  labs(x = "Percentage of students in school", y = "Count")+
  theme(plot.title = element_text(face="bold", size=14, hjust=0))+
  theme(axis.title = element_text(face="bold", size=14))+ 
  theme(axis.text.x  = element_text(size=14),axis.text.y  = element_text(size=14))+
  theme(legend.text=element_text(size=14),legend.title=element_text(size=14))+
  theme(text=element_text(family="Times"))+
  theme(legend.position = "bottom")+
  theme(strip.text = element_text(size=14))+
  scale_x_continuous(labels=percent)+
  scale_fill_manual(values = c("#bebad9", "#3b6dae"), name="Chronic absenteeism", breaks=c("0","1"), labels=c("No", "Yes"))
```

\newpage